Proofs for ``Geometry of Local Spectral Clustering''

### 2021-01-24 jmlr/revision/proofs.tex
All proofs completed.

The proofs of our major theorems largely consist of (at most) three modular parts.
\begin{enumerate}
	\item \textbf{Fixed graph results.} Results which hold with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item \textbf{Finite-sample bounds.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item \textbf{Population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\paragraph{Organization.} In the following subsections we establish: (\ref{subsec:pf_lem_zhu}) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}); (\ref{subsec:ppr_uniform_bounds}) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR);  (\ref{subsec:lovasz_simonovits_bounds}) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (\ref{subsec:pf_prop_pointwise_mixing_time}) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); and (\ref{subsec:ppr_spectral_partitioning}) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example).

\subsection{Proof of Lemma~\ref{lem:zhu}}
\label{subsec:pf_lem_zhu}
For a candidate cluster $C \subseteq V$, we use the tilde-notation $\wt{G} = G[C]$ to refer to the subgraph of $G$ induced by $C$. Similarly we write $\wt{q}_v^{(t)} := q(v,t;\wt{G})$ for the $t$-step distribution of the lazy random walk over $\wt{G}$, $\wt{\pi} = \pi(G[C])$ for the stationary distribution of $\wt{q}_v^{(t)}$ (we will always assume $G[C]$ is connected), and $\wt{p}_v := p(v,\alpha;\wt{G})$ for the PPR vector over $\wt{G}$.

As mentioned in the main text, Lemma~\ref{lem:zhu} is equivalent, up to constants, to Lemma~3.4 in \cite{zhu2013}, and the proof of Lemma~\ref{lem:zhu} proceeds along very similar lines to the proof of that lemma. In fact, we directly use the following three inequalities, derived in that work:
\begin{itemize}
	\item \textbf{(c.f. Lemma 3.2 of \cite{zhu2013})} For any seed node $v \in C$, the PPR vector is lower bounded,
	\begin{equation}
	\label{pf:zhu1}
	\wt{p}_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Corollary 3.3 of \cite{zhu2013})} For any seed node $v \in C$, there exists a so-called leakage distribution $\ell = \ell(v)$ such that $\mathrm{supp}(\ell) \subseteq C$, $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$, and 
	\begin{equation}
	\label{pf:zhu2}
	p_v(u) \geq \wt{p}_v(u) - \wt{p}_{\ell}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Lemma 3.1 of \cite{zhu2013})} There exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C;G)$ such that for any seed node $v \in C^g$, the following inequality holds
	\begin{equation}
	\label{pf:zhu3}
	p_v(C^c) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation}
\end{itemize}
We use~\eqref{pf:zhu1}-\eqref{pf:zhu3} to separately upper bound $\vol(S_{\beta,v} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$ and $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$; here $C^{\mathrm{int}} \cup C^{\mathrm{bdry}} = C$ is a partition of $C$, with
\begin{equation*}
C^{\mathrm{int}} := \Bigl\{u \in C: \deg(u;\wt{G}) > \bigl(1 - \alpha \cdot \beta \cdot \vol(C;G)\bigr) \deg(u;G) \Bigr\},
\end{equation*}
consisting of those vertices $u \in C$ with sufficient large degree in $\wt{G}$.

@@

Next we upper bound $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$. From~\eqref{pf:zhu1} and~\eqref{pf:zhu2} we see that 
\begin{equation*}
p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$.}
\end{equation*}
If additionally $u \not\in S_{\beta,v}$ then $p_v(u) \leq \beta \deg(u;G)$, and for all such $u \in C \setminus S_{\beta,v}$,
\begin{equation}
\label{pf:zhu4}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  \beta\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
On the other hand, for any $u \in C^{\mathrm{int}}$ it holds that
\begin{equation*}
\wt{\pi}(u) = \frac{\deg(u;\wt{G})}{\vol(\wt{G})} \geq \frac{\deg(u;\wt{G})}{\vol(G)} \geq \frac{(1 - \alpha \beta \vol(C;G))\deg(u;G)}{\vol(C;G)};
\end{equation*}
by plugging this in to~\eqref{pf:zhu4} we obtain
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \deg(u;G) \leq \wt{p}_{\ell}(u),~~\textrm{for all $u \in C^{\mathrm{int}} \setminus S_{\beta,v}$},
\end{equation*}
and summing over all such $u$ gives
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds in~\eqref{eqn:zhu_condition} imply
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu5}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}

Finally, we upper bound $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$. Indeed, for any $u \in C^{\mathrm{bdry}}$,
\begin{equation*}
\frac{1}{\vol(C;G)}\sum_{w \not\in C} \1((u,w) \in E) \geq \alpha \cdot \beta \cdot \deg(u;G)
\end{equation*}
and summing over all such vertices yields
\begin{equation}
\label{pf:zhu6}
\vol(C^{\mathrm{bdry}};G) \leq \frac{1}{\alpha \beta \vol(C;G)}\sum_{\substack{u \in C^{\mathrm{bdry}} \\ w \not\in C}} \1((u,w) \in E) \leq \frac{\Phi(C;G)}{\alpha \cdot \beta}.
\end{equation} 
The claim follows upon summing the upper bounds in~\eqref{pf:zhu3.5}, \eqref{pf:zhu5} and~\eqref{pf:zhu6}.

@@

\subsection{Sample-to-population: normalized cut}
\label{subsec:sample_to_population_ncut}
In this subsection we establish~\eqref{eqn:sample_to_population_normalized_cut}. For a set $\mc{S} \subseteq \Rd$, both $\mathrm{cut}_{n,r}(\mc{S}[X])$ and $\mathrm{vol}_{n,r}(\mc{S}[X])$ are order-$2$ U-statistics:
\begin{align*}
\mathrm{cut}_{n,r}(\mc{S}[X]) & = \sum_{i = 1}^{n} \sum_{j \neq i} \1\{\|x_i - x_j\| \leq r\} \cdot \1\{x_i \in \mc{S}\} \cdot \1\{x_j \not\in \mc{S}\},
\intertext{and}
\mathrm{vol}_{n,r}(\mc{S}[X]) & = \sum_{i = 1}^{n} \sum_{j \neq i} \1\{\|x_i - x_j\| \leq r\} \cdot \1\{x_i \in \mc{S}\}.
\end{align*}
Therefore with probability at least $1 - \exp\{-\delta^2 \cut_{\Pbb,r}(\mc{S})\}$,
\begin{equation*}
\frac{1}{n(n - 1)} \mathrm{cut}_{n,r}(\mc{S}[X]) \leq (1 + \delta) \cut_{\Pbb,r}(\mc{S}),
\end{equation*}
and likewise with probability at least $1 - \exp\{-\delta^2 \vol_{\Pbb,r}(\mc{S})\}$,
\begin{equation*}
(1 - \delta) \vol_{\Pbb,r}(\mc{S}) \leq \frac{1}{n(n - 1)} \vol_{n,r}(\mc{S}[X]).
\end{equation*}
Consequently, for any $\delta \in (0,1/3)$,  
\begin{equation*}
\Phi_{n,r}(\mc{C}[X]) \leq \frac{1 + \delta}{1 - \delta} \frac{\cut_{\Pbb,r}(\mc{C})}{\min\{\vol_{\Pbb,r}(\mc{C}), \vol_{\Pbb,r}(\mc{C}^c)\}} = \frac{1 + \delta}{1 - \delta} \Phi_{\Pbb,r}(\mc{C}) \leq (1 + 3\delta) \cdot \Phi_{\Pbb,r}(\mc{C})  
\end{equation*}
with probability at least $1 - 3\exp\{-\delta^2 \cut_{\Pbb,r}(\mc{C})\}$. This establishes~\eqref{eqn:sample_to_population_normalized_cut} upon taking $b_1 = \cut_{\Pbb,r}(\mc{C})$ and $C_1 = 3$.

\subsection{Sample-to-population: local spread}
\label{subsec:sample_to_population_local_spread}
In this subsection we establish~\eqref{eqn:sample_to_population_local_spread}. To ease the notational burden, let $\wt{G}_{n,r} := G_{n,r}\bigl[\mc{C}[X]\bigr]$. By Lemma~\ref{lem:hoeffding}, with probability at least $1 - \exp\{-2\delta^2 \Pbb(\mc{C})^2n\}$
\begin{equation}
\label{pf:sample_to_population_local_spread_1}
\wt{n} \geq (1 - \delta) \cdot n \cdot \Pbb(\mc{C}).
\end{equation}
Conditional on~\eqref{pf:sample_to_population_local_spread_1}, it follows from Lemma~\ref{lem:bernstein_union} that with probability at least $1 -n\exp\{-2(1 - \delta)\delta^2d_{\min}(\wt{\Pbb}) \Pbb(\mc{C}) n\}$,
\begin{equation}
\label{pf:sample_to_population_local_spread_2}
\frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq (1 - \delta) \cdot d_{\min}(\wt{\Pbb}),
\end{equation}
and likewise it follows from Lemma~\ref{lem:hoeffding_2} that with probability at least $1 - \exp\{-2\delta^2(1 - \delta) \vol_{\wt{\Pbb},r}(\mc{C})^2 \Pbb(\mc{C})n \}$,
\begin{equation*}
\frac{1}{\wt{n}(\wt{n} - 1)} \vol(\wt{G}_{n,r}) \geq (1 - \delta) \cdot \vol_{\wt{\Pbb},r}(\mc{C}).
\end{equation*}
Consequently, for any $\delta \in (0,1/3)$,
\begin{equation*}
s_{n,r}(\mc{C}[X]) = \frac{d_{\min}(\wt{G}_{n,r})^2}{\vol(\wt{G}_{n,r})} \geq \frac{\frac{1}{\wt{n}^2}d_{\min}(\wt{G}_{n,r})^2}{\frac{1}{\wt{n}(\wt{n} - 1)}\vol(\wt{G}_{n,r})} \geq \frac{(1 - \delta)^2}{(1 + \delta)} \frac{d_{\min}(\wt{\Pbb})^2}{\vol_{\wt{\Pbb},r}(\mc{C})} \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{C}).
\end{equation*}
with probability at least $1 - 2\exp\bigl\{-\delta^2\bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})n\bigr\} - n\exp\bigl\{-\delta^2\bigl(d_{\min}(\wt{\Pbb})\bigr)^2 \Pbb(\mc{C}) n\bigr\}$. This establishes~\eqref{eqn:sample_to_population_local_spread} upon taking $b_2 = \bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})$ and $b_3 = d_{\min}(\wt{\Pbb}) \Pbb(\mc{C})$

@@

		\item \textbf{Volume of $\delta$-interiors.} Suppose $\nu(\mc{S}^{\delta}) \leq (1 - a) \cdot \nu(\mc{S})$ or $\nu(\mc{R}^{\delta}) \leq (1 - a) \cdot \nu(\mc{R})$. Then 

@@

\section{Proof of Major Theorems}
\label{sec:pf_major_theorems}
In this section, we prove Theorem~\ref{thm:volume_ssd_ub} (in Section), Corollary~\ref{cor:density_cluster_volume_ssd_ub}, and Theorem~\ref{thm:ppr_lb}.

\subsection{Proof of Theorem~\ref{thm:volume_ssd_ub}}

\subsection{Proof of Corollary~\ref{cor:density_cluster_volume_ssd_ub}}

>>>>

The proofs of our major theorems largely consist of (at most) three modular parts.
\begin{enumerate}
	\item \textbf{Fixed graph results.} Results which hold with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item \textbf{Sample-to-population results.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item \textbf{Bounds on population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\paragraph{Organization.} In the following subsections we establish: (\ref{subsec:pf_lem_zhu}) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}), and an analogous result for aPPR (Corollary~\ref{cor:zhu}); (\ref{subsec:ppr_uniform_bounds}) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR);  (\ref{subsec:lovasz_simonovits_bounds}) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (\ref{subsec:pf_prop_pointwise_mixing_time}) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); and (\ref{subsec:ppr_spectral_partitioning}) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example).

\subsection{Misclassification error of clustering with PPR and aPPR}
\label{subsec:pf_lem_zhu}
For a candidate cluster $C \subseteq V$, we use the tilde-notation $\wt{G} = G[C]$ to refer to the subgraph of $G$ induced by $C$. Similarly we write $\wt{q}_v^{(t)} := q(v,t;\wt{G})$ for the $t$-step distribution of the lazy random walk over $\wt{G}$, $\wt{\pi} = \pi(G[C])$ for the stationary distribution of $\wt{q}_v^{(t)}$ (we will always assume $G[C]$ is connected), and $\wt{p}_v := p(v,\alpha;\wt{G})$ for the PPR vector over $\wt{G}$.

\begin{proof}[Proof (of Lemma~\ref{lem:zhu})]
As mentioned in the main text, Lemma~\ref{lem:zhu} is equivalent, up to constants, to Lemma~3.4 in \cite{zhu2013}, and the proof of Lemma~\ref{lem:zhu} proceeds along very similar lines to the proof of that lemma. In fact, we directly use the following three inequalities, derived in that work:
\begin{itemize}
	\item \textbf{(c.f. Lemma 3.2 of \cite{zhu2013})} For any seed node $v \in C$, the PPR vector is lower bounded,
	\begin{equation}
	\label{pf:zhu1}
	\wt{p}_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Corollary 3.3 of \cite{zhu2013})} For any seed node $v \in C$, there exists a so-called leakage distribution $\ell = \ell(v)$ such that $\mathrm{supp}(\ell) \subseteq C$, $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$, and 
	\begin{equation}
	\label{pf:zhu2}
	p_v(u) \geq \wt{p}_v(u) - \wt{p}_{\ell}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Lemma 3.1 of \cite{zhu2013})} There exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C;G)$ such that for any seed node $v \in C^g$, the following inequality holds
	\begin{equation}
	\label{pf:zhu3}
	p_v(C^c) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation}
\end{itemize}
We use~\eqref{pf:zhu1}-\eqref{pf:zhu3} to separately upper bound $\vol(S_{\beta,v} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$ and $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$; here $C^{\mathrm{int}} \cup C^{\mathrm{bdry}} = C$ is a partition of $C$, with
\begin{equation*}
C^{\mathrm{int}} := \Bigl\{u \in C: \deg(u;\wt{G}) > \bigl(1 - \alpha \cdot \beta \cdot \vol(C;G)\bigr) \deg(u;G) \Bigr\},
\end{equation*}
consisting of those vertices $u \in C$ with sufficient large degree in $\wt{G}$.

@@

Next we upper bound $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$. From~\eqref{pf:zhu1} and~\eqref{pf:zhu2} we see that 
\begin{equation*}
p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$.}
\end{equation*}
If additionally $u \not\in S_{\beta,v}$ then $p_v(u) \leq \beta \deg(u;G)$, and for all such $u \in C \setminus S_{\beta,v}$,
\begin{equation}
\label{pf:zhu4}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  \beta\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
On the other hand, for any $u \in C^{\mathrm{int}}$ it holds that
\begin{equation*}
\wt{\pi}(u) = \frac{\deg(u;\wt{G})}{\vol(\wt{G})} \geq \frac{\deg(u;\wt{G})}{\vol(G)} \geq \frac{(1 - \alpha \beta \vol(C;G))\deg(u;G)}{\vol(C;G)};
\end{equation*}
by plugging this in to~\eqref{pf:zhu4} we obtain
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \deg(u;G) \leq \wt{p}_{\ell}(u),~~\textrm{for all $u \in C^{\mathrm{int}} \setminus S_{\beta,v}$};
\end{equation*}
and summing over all such $u$ gives
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds on $\alpha$ and $\beta$ in~\eqref{eqn:zhu_condition} imply
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu5}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}

Finally, we upper bound $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$. Indeed, for any $u \in C^{\mathrm{bdry}}$,
\begin{equation*}
\frac{1}{\vol(C;G)}\sum_{w \not\in C} \1((u,w) \in E) \geq \alpha \cdot \beta \cdot \deg(u;G)
\end{equation*}
and summing over all such vertices yields
\begin{equation}
\label{pf:zhu6}
\vol(C^{\mathrm{bdry}};G) \leq \frac{1}{\alpha \beta \vol(C;G)}\sum_{\substack{u \in C^{\mathrm{bdry}} \\ w \not\in C}} \1((u,w) \in E) \leq \frac{\Phi(C;G)}{\alpha \cdot \beta}.
\end{equation} 
The claim follows upon summing the upper bounds in~\eqref{pf:zhu3.5}, \eqref{pf:zhu5} and~\eqref{pf:zhu6}.
\end{proof}

If the cluster estimate $\wh{C}$ is instead obtained by sweep cutting the aPPR vector $p_v^{(\varepsilon)}$, a similar upper bound on $\vol(\wh{C} \vartriangle C)$ holds, provided that $\varepsilon$ is sufficiently small.
\begin{corollary}
	\label{cor:zhu}
	For a set $C \subseteq V$, suppose that $\alpha,\beta$ satisfy~\eqref{eqn:zhu_condition}, and additionally that 
	\begin{equation}
	\label{eqn:zhu_condition_2}
	\varepsilon \leq \frac{1}{25\vol(C;G)}.
	\end{equation}
	Then there exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C^g;G)$ such that for any $v \in C^g$, the sweep cut $S_{\beta,v}$ of the aPPR vector $p_v^{(\varepsilon)}$ satisfies
	\begin{equation}
	\label{eqn:zhu_ub}
	\vol(S_{\beta,v} \vartriangle C;G) \leq 6\frac{\Phi(C;G)}{\alpha \beta}.
	\end{equation}
\end{corollary}
\begin{proof}[Proof (of Corollary~\ref{cor:zhu})]
	Recall that the upper bound~\eqref{eqn:zhu_ub} on $\vol(\wh{C} \vartriangle C;G)$ comes from combining the upper bounds on $\vol(\wh{C} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus \wh{C};G)$ and $\vol(C^{\mathrm{bdry}} \setminus \wh{C};G)$ in~\eqref{pf:zhu3.5},~\eqref{pf:zhu5} and~\eqref{pf:zhu6}. From the upper bound $p_v^{(\varepsilon)}(u) \leq p_v(u)$ for all $u \in V$, it is clear that both~\eqref{pf:zhu3.5} and~\eqref{pf:zhu6} continue to hold when the aPPR vector is used instead of the PPR vector.

	It remains only to establish an upper bound on~$\vol(C^{\mathrm{int}} \setminus \wh{C};G)$. For any $u \in C \setminus S_{\beta,v}$, from inequality~\eqref{pf:zhu2} and the lower bound $p_v^{(\varepsilon)}(u) \geq p_v(u)  - \varepsilon \deg(u;G)$ in~\eqref{eqn:appr_error} we deduce that
	\begin{equation}
	\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  (\beta + \varepsilon)\deg(u;G) \leq \wt{p}_{\ell}(u).
	\end{equation}
	Following the same steps as used in the proof of Lemma~\ref{lem:zhu} yields the following inequality:
	\begin{equation*}
	\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation*}
	The upper bounds on $\alpha,\beta$ in~\eqref{eqn:zhu_condition},  and on $\varepsilon$ in~\eqref{eqn:zhu_condition_2}, imply that
	\begin{equation*}
	\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
	\end{equation*}
	and we conclude that
	\begin{equation}
	\label{pf:zhu2_1}
	\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
	\end{equation}
	Summing the right hand sides of~\eqref{pf:zhu2},~\eqref{pf:zhu6}, and~\eqref{pf:zhu2_1} yields the claim.
\end{proof}

@@

\subsection{Sample-to-population: normalized cut}
\label{subsec:sample_to_population_ncut}
In this subsection we establish~\eqref{eqn:sample_to_population_normalized_cut}. For a set $\mc{S} \subseteq \Rd$, both $\mathrm{cut}_{n,r}(\mc{S}[X])$ and $\mathrm{vol}_{n,r}(\mc{S}[X])$ are order-$2$ U-statistics:
\begin{align*}
\mathrm{cut}_{n,r}(\mc{S}[X]) & = \sum_{i = 1}^{n} \sum_{j \neq i} \1\{\|x_i - x_j\| \leq r\} \cdot \1\{x_i \in \mc{S}\} \cdot \1\{x_j \not\in \mc{S}\},
\intertext{and}
\mathrm{vol}_{n,r}(\mc{S}[X]) & = \sum_{i = 1}^{n} \sum_{j \neq i} \1\{\|x_i - x_j\| \leq r\} \cdot \1\{x_i \in \mc{S}\}.
\end{align*}
Therefore with probability at least $1 - \exp\{-\delta^2 (\cut_{\Pbb,r}(\mc{S}))^2\}$,
\begin{equation*}
\frac{1}{n(n - 1)} \mathrm{cut}_{n,r}(\mc{S}[X]) \leq (1 + \delta) \cut_{\Pbb,r}(\mc{S}),
\end{equation*}
and likewise with probability at least $1 - \exp\{-\delta^2 (\vol_{\Pbb,r}(\mc{S}))^2\}$,
\begin{equation*}
(1 - \delta) \vol_{\Pbb,r}(\mc{S}) \leq \frac{1}{n(n - 1)} \vol_{n,r}(\mc{S}[X]).
\end{equation*}
Consequently, for any $\delta \in (0,1/3)$,  
\begin{equation*}
\Phi_{n,r}(\mc{C}[X]) \leq \frac{1 + \delta}{1 - \delta} \cdot  \frac{\cut_{\Pbb,r}(\mc{C})}{\min\{\vol_{\Pbb,r}(\mc{C}), \vol_{\Pbb,r}(\mc{C}^c)\}} = \frac{1 + \delta}{1 - \delta} \cdot \Phi_{\Pbb,r}(\mc{C}) \leq (1 + 3\delta) \cdot \Phi_{\Pbb,r}(\mc{C})  
\end{equation*}
with probability at least $1 - 3\exp\{-\delta^2 (\cut_{\Pbb,r}(\mc{C}))^2\}$. This establishes~\eqref{eqn:sample_to_population_normalized_cut} upon taking $b_1 := (\cut_{\Pbb,r}(\mc{C}))^2$.

\subsection{Sample-to-population: local spread}
\label{subsec:sample_to_population_local_spread}
In this subsection we establish~\eqref{eqn:sample_to_population_local_spread}. To ease the notational burden, let $\wt{G}_{n,r} := G_{n,r}\bigl[\mc{C}[X]\bigr]$. By Lemma~\ref{lem:hoeffding_2}, with probability at least $1 - \exp\{-2\delta^2 (\Pbb(\mc{C}))^2n\}$
\begin{equation}
\label{pf:sample_to_population_local_spread_1}
\wt{n} \geq (1 - \delta) \cdot n \cdot \Pbb(\mc{C}).
\end{equation}
Conditional on~\eqref{pf:sample_to_population_local_spread_1}, it follows from Lemma~\ref{lem:bernstein_union} that with probability at least $1 -n\exp\{-\delta^2d_{\min}(\wt{\Pbb}) \cdot (1 - \delta) \Pbb(\mc{C}) n/(3 + \delta)\}$,
\begin{equation}
\label{pf:sample_to_population_local_spread_2}
\frac{1}{\wt{n}} d_{\min}(\wt{G}_{n,r}) \geq (1 - \delta) \cdot d_{\min}(\wt{\Pbb}),
\end{equation}
and likewise it follows from Lemma~\ref{lem:hoeffding} that with probability at least $1 - \exp\{-\delta^2(\vol_{\wt{\Pbb},r}(\mc{C}))^2 \cdot (1 - \delta) \Pbb(\mc{C}) n \}$,
\begin{equation*}
\frac{1}{\wt{n}(\wt{n} - 1)} \vol(\wt{G}_{n,r}) \geq (1 - \delta) \cdot \vol_{\wt{\Pbb},r}(\mc{C}).
\end{equation*}
Consequently for any $\delta \in (0,1/3)$,
\begin{equation*}
s_{n,r}(\mc{C}[X]) = \frac{d_{\min}(\wt{G}_{n,r})^2}{\vol(\wt{G}_{n,r})} \geq \frac{\frac{1}{\wt{n}^2}d_{\min}(\wt{G}_{n,r})^2}{\frac{1}{\wt{n}(\wt{n} - 1)}\vol(\wt{G}_{n,r})} \geq \frac{(1 - \delta)^2}{(1 + \delta)} \cdot  \frac{d_{\min}(\wt{\Pbb})^2}{\vol_{\wt{\Pbb},r}(\mc{C})} \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{C}).
\end{equation*}
with probability at least $1 - \exp\{-2\delta^2 (\Pbb(\mc{C}))^2n\} - \exp\bigl\{-2\delta^2\bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})n/3\bigr\} - n\exp\bigl\{-\delta^2 d_{\min}(\wt{\Pbb}) \cdot \Pbb(\mc{C}) n/5\bigr\}$. This establishes~\eqref{eqn:sample_to_population_local_spread} upon taking $b_2 := (\Pbb(\mc{C}))^2 \wedge 2\bigl(\vol_{\wt{\Pbb},r}(\mc{C})\bigr)^2 \Pbb(\mc{C})/3 \wedge d_{\min}(\wt{\Pbb}) \cdot \Pbb(\mc{C})/5$.

@@

		\item \textbf{Lebesgue measure of $\delta$-interiors.} Suppose $\nu(\mc{S}^{\delta}) \leq (1 - a) \cdot \nu(\mc{S})$ or $\nu(\mc{R}^{\delta}) \leq (1 - a) \cdot \nu(\mc{R})$. Then 

@@

\section{Proof of Major Theorems}
\label{sec:pf_major_theorems}
We now prove the three major theorems of our paper: Theorem~\ref{thm:volume_ssd_ub} (in Section~\ref{subsec:pf_volume_ssd_ub}), Theorem~\ref{thm:density_cluster_volume_ssd_ub}, and Theorem~\ref{thm:ppr_lb}.  Throughout, we use the notation $\wt{n} = |\mc{C}[X]|$ and $\wt{G}_{n,r} = G_{n,r}\bigl[\mc{C}[X]\bigr]$ as defined above.
